---
title: "differential genes in lupus single cell seq"
author: "sara yusuf"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---







```{r}
library(Seurat , verbose = F)
library(tidyverse, verbose = F)
library(ExperimentHub, verbose = F)
library(metap, verbose = F)
library(multtest, verbose = F)
library(SingleCellExperiment, verbose = F)
library(DESeq2)
library(MAST)
```


```{r}
# get data
eh <- ExperimentHub()
query(eh, "Kang")

sce <- eh[["EH2259"]]
seu.obj <- as.Seurat(sce, data = NULL)

head(seu.obj@meta.data)

```
```{r}
str(seu.obj)
```





```{r}
# get mito percent
seu.obj$mitoPercent <- PercentageFeatureSet(seu.obj, pattern = '^MT-')
head(seu.obj@meta.data)

# filter
seu.filtered <- subset(seu.obj, subset = nFeature_originalexp > 200 & nFeature_originalexp < 2500 &
         nCount_originalexp > 800 & 
         mitoPercent < 5 &
         multiplets == 'singlet')

seu.obj
seu.filtered
```

Purpose of FindVariableFeatures()

1. Identify Key Genes:

High Variability: The function helps identify genes with high variability in their expression levels across cells. These genes are often more useful for distinguishing biological differences between cells.
Dimensionality Reduction: By focusing on these highly variable genes, you can reduce the dimensionality of the data, making it easier to perform clustering and visualization.

2. Improve Clustering and Visualization:

Clustering: Highly variable genes provide better input for clustering algorithms because they capture the most significant biological differences between cells.
Visualization: Techniques like PCA (Principal Component Analysis) and t-SNE (t-Distributed Stochastic Neighbor Embedding) or UMAP (Uniform Manifold Approximation and Projection) benefit from focusing on variable genes for clearer separation of clusters.


```{r}
seu.filtered <- NormalizeData(seu.filtered)
seu.filtered <- FindVariableFeatures(seu.filtered)
seu.filtered <- ScaleData(seu.filtered)
seu.filtered <- RunPCA(seu.filtered)
```

```{r}
str(seu.filtered)
```


```{r}
variablefeatures = VariableFeatures(seu.filtered)
length(variablefeatures)
```


```{r}
# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(seu.filtered), 10)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(seu.filtered)
LabelPoints(plot = plot1, points = top10, repel = TRUE)


```



```{r}
ElbowPlot(seu.filtered)

```


```{r}
seu.filtered <- RunUMAP(seu.filtered, dims = 1:20)

```


```{r}
# visualize 
cell_plot <- DimPlot(seu.filtered, reduction = 'umap', group.by = 'cell', label = TRUE)
cond_plot <- DimPlot(seu.filtered, reduction = 'umap', group.by = 'stim')

cell_plot|cond_plot
```


Purpose of FindAllMarkers()

1. Identify Cluster-Specific Markers:

Cluster Markers: Helps in finding genes that are specifically expressed in individual clusters, which can be used to characterize and distinguish different cell types or states.
Comparative Analysis: Assesses differential expression across all clusters, providing a comprehensive view of which genes are unique to each cluster.

2. Understand Cellular Heterogeneity:

Biological Insight: Provides insight into the molecular signatures of different cell populations, aiding in the interpretation of cellular heterogeneity.

```{r}

Idents(seu.filtered) = seu.filtered@meta.data$cell

```


```{r}
# findAll markers 

#FindAllMarkers(seu.filtered,
             # logfc.threshold = 0.25,
             #  min.pct = 0.1,
             # only.pos = TRUE,
             #  test.use = 'DESeq2',
             #  slot = 'counts')

```


```{r}
#save the variable 
#save(all_markers, file = "all_markers.RData")


#Loading the Variable from a File
load("all_markers.RData")
```


```{r}
# Identify top 10 markers for each cluster
top10_all_markers <- all_markers %>% 
  group_by(cluster) %>% 
  top_n(n = 10, wt = avg_log2FC)




# Plot heatmap 
DoHeatmap(seu.filtered, features = unique(top10_all_markers$gene)) + 
  NoLegend() + 
  theme(legend.position = "bottom") + 
  scale_fill_gradientn(colors = c("blue", "white", "red")) +
  theme(legend.text = element_text(size = 5))

```
```{r}
top10_all_markers <- top10_all_markers[!duplicated(top10_all_markers$gene), ]

DotPlot(
  seu.filtered,
  top10_all_markers$gene,
  assay = NULL,
  cols = c("lightgrey", "red"),
  col.min = -2.5,
  col.max = 2.5,
  dot.min = 0,
  dot.scale = 6,
  idents = NULL,
  group.by = NULL,
  split.by = NULL,
  cluster.idents = FALSE,
  scale = TRUE,
  scale.by = "radius",
  scale.min = NA,
  scale.max = NA
) + theme(axis.text.x = element_text(angle = 90, hjust = 1 , size = 5))

```




```{r}
#Feature plot for top expressed gebes in each cell type genes

FeaturePlot(seu.filtered, features = c("CD83", "NKG7" ,"HIST1H2AC","TXN"  ) , label = T)
```



```{r}
DefaultAssay(seu.filtered) 


```


Purpose of FindConservedMarkers()

1. Identify Stable Markers:

Consistency Across Conditions: This function helps find markers that are consistently expressed in certain cell types or clusters regardless of experimental conditions or different datasets.
Robust Features: Conserved markers are useful for understanding the core characteristics of cell types or states that are preserved across different biological contexts.

2. Cross-Dataset Comparisons:

Benchmarking: Allows comparison of markers across different datasets or experiments, which is valuable for validating findings and ensuring the robustness of identified markers.

```{r}
# findConserved markers for "CD14+ Monocytes"

#markers_cluster <- FindConservedMarkers(seu.filtered,
#                    ident.1 = "CD14+ Monocytes",
#                     grouping.var = 'stim',
#                      assay  =  "originalexp")
                     
```


```{r}

#save the variable 
#save(markers_cluster, file = "markers_cluster.RData")

#Loading the Variable from a File
load("markers_cluster.RData")
head(markers_cluster)

```

```{r}
head(markers_cluster)
```

```{r}

library(dplyr)

filtered_con <- markers_cluster %>%
  filter(minimump_p_val < 0.05 ) %>%
  slice_head(n = 100)


```


```{r}
Idents(seu.filtered) = seu.filtered$cell
```



```{r}
DotPlot(
  seu.filtered,
  rownames(filtered_con),
  assay = NULL,
  cols = c("lightgrey", "red"),
  col.min = -2.5,
  col.max = 2.5,
  dot.min = 0,
  dot.scale = 6,
  idents = NULL,
  group.by = NULL,
  split.by = NULL,
  cluster.idents = FALSE,
  scale = TRUE,
  scale.by = "radius",
  scale.min = NA,
  scale.max = NA
) + theme(axis.text.x = element_text(angle = 90, hjust = 1 , size = 5))

```



#visualize "TYROBP" gene 

```{r}
cell_plot <- DimPlot(seu.filtered, reduction = 'umap', group.by = 'cell', label = TRUE)
cell_plot
```


```{r}
FeaturePlot(seu.filtered, features = 'TYROBP', min.cutoff = 'q10')

```




```{r}
# findMarkers between conditions 

seu.filtered$celltype.cnd <- paste0(seu.filtered$cell,"_" ,seu.filtered$stim)
Idents(seu.filtered) <- seu.filtered$celltype.cnd
head(seu.filtered@meta.data)

```


```{r}
DimPlot(seu.filtered, reduction = 'umap', label = TRUE)

```



 find markers differentiate CD14+ Monocytes_stim vs CD14+ Monocytes_ctrl using all listed methods

```{r}

# Define the list of test methods to compare
#test_methods <- c("wilcox", "bimod", "t", "LR", "MAST", "DESeq2")

# Create an empty list to store the results for each test
#results_list <- list()

#Loop through each test method and run FindMarkers
#   for (test in test_methods) {
     #Run FindMarkers with the current test method
 #    markers <- FindMarkers(seu.filtered, 
#                        ident.1 = 'CD14+ Monocytes_stim', 
 #                       ident.2 = 'CD14+ Monocytes_ctrl', 
 #                      test.use = test)
  
  # Store the results in the list with the test method as the name
# results_list[[test]] <- markers
#}


```




```{r}
#save(markers, file = "markers.RData")

#Loading the Variable from a File
load("markers.RData")

```

```{r}
load("results_list.RData")

```




```{r}



# Loop through each element in markers_list and assign to a new data frame
for (test in names(results_list)) {
  assign(paste0("DEGs_", test), results_list[[test]])
}

```


```{r}
library(dplyr)

# Combine all DEGs into one data frame with a method column
combined <- do.call(rbind, lapply(names(results_list), function(test) {
  df <- results_list[[test]]
  df$method <- test  # Add method name as a new column
  df
}))


library(ggplot2)

# Choose a refined color palette
color_palette <- c("#2E86C1", "#E74C3C", "#27AE60", "#8E44AD", "#F39C12", "#16A085", "#D35400", "#7F8C8D")

# Create the box plot with enhanced aesthetics
ggplot(combined, aes(x = method, y = -log10(p_val_adj), color = method)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.8, width = 0.6, lwd = 0.6) +  # Slightly thicker lines for boxes
  geom_jitter(width = 0.2, size = 0.8, alpha = 0.6) +  # Smaller, more subtle jittered points
  labs(title = "Distribution of Adjusted P-Values Across Methods",
       subtitle = "A comparison of adjusted p-values for DEGs across different analysis methods",
       x = "Method",
       y = "Adjusted P-Value",
       caption = "Log10 scale used for p-values. Each point represents a gene's adjusted p-value.") +
  theme_classic(base_size = 14) +  # Classic theme with a base font size suitable for publication
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, face = "bold", size = 12),  # Bold x-axis labels
    axis.text.y = element_text(face = "bold", size = 12),  # Bold y-axis labels
    plot.title = element_text(hjust = 0.5, face = "bold", size = 16),  # Centered and bold title
    plot.subtitle = element_text(hjust = 0.5, face = "italic", size = 12),  # Centered italic subtitle
    legend.position = "none",  # Remove legend to avoid redundancy
    panel.grid.major.y = element_line(color = "gray90", size = 0.3)  # Light grid lines for y-axis
  ) +
  scale_y_log10(labels = scales::scientific) +  # Log scale with scientific notation for p-values
  scale_color_manual(values = color_palette)  # Apply the refined color palette


```



```{r}
filter_degs <- function(df) {
  df %>%
    filter(p_val_adj < 0.05 & avg_log2FC > 2)
}

# Apply the filtering function to each data frame
sig_wilcox <- filter_degs(DEGs_wilcox)
sig_bimod <- filter_degs(DEGs_bimod)
sig_t <- filter_degs(DEGs_t)
sig_LR <- filter_degs(DEGs_LR)
sig_MAST <- filter_degs(DEGs_MAST)
sig_DESeq2 <- filter_degs(DEGs_DESeq2)
```



```{r}
# List of significant markers
sig_genes_list <- list(
  "wilcox" = sig_wilcox,
  "bimod" = sig_bimod,
  "t" = sig_t,
  "LR" = sig_LR,
  "MAST" = sig_MAST,
  "DESeq2" = sig_DESeq2
)

```








```{r}
# Loop through each element in sig_genes_list and write to a CSV file
#for (list_name in names(results_list)) {
  # Assign each list element to a data frame
# df <- results_list[[list_name]]
  
 #Write the data frame to a CSV file
 # write.csv(df, file = paste0("sig_", list_name, ".csv"), row.names = TRUE , sep = "")
#}

```

```{r}
#sig_bimod  <- read.csv("C:/Users/saraa/Desktop/sig_bimod.csv" , row.names  = 1, header = T)
#sig_DESeq2 <- read.csv("C:/Users/saraa/Desktop/sig_DESeq2.csv", row.names  = 1, header = T)
#sig_LR     <- read.csv("C:/Users/saraa/Desktop/sig_LR.csv", row.names  = 1, header = T) 
#sig_MAST   <- read.csv("C:/Users/saraa/Desktop/sig_MAST.csv", row.names  = 1, header = T)
#sig_t      <- read.csv("C:/Users/saraa/Desktop/sig_t.csv.", row.names  = 1, header = T)
#sig_wilcox <- read.csv("C:/Users/saraa/Desktop/sig_wilcox.csv", row.names  = 1, header = T)
```


```{r}
# Get the rownames for each data frame in the list
row_names_list <- lapply(sig_genes_list, rownames)

# Find the common rownames across all data frames
common_genes <- Reduce(intersect, row_names_list)

# Print the common genes
length(common_genes)

```






```{r}
library(dplyr)
library(ggplot2)

# Combine all significant dataframes into one, adding a method column
combined_df <- bind_rows(
  mutate(DEGs_wilcox, method = "Wilcox"),
  mutate(DEGs_bimod, method = "Bimod"),
  mutate(DEGs_t, method = "T"),
  mutate(DEGs_LR, method = "LR"),
  mutate(DEGs_MAST, method = "MAST"),
  mutate(DEGs_DESeq2, method = "DESeq2")
)

# Add a column for color based on the fold change and significance threshold
combined_df <- combined_df %>%
  mutate(color = case_when(
    p_val_adj < 0.05 & avg_log2FC > 2 ~ "up",
    p_val_adj > 0.05 & abs(avg_log2FC) <= 2 ~ "Not Significant",
    p_val_adj < 0.05 & avg_log2FC <= -2 ~ "down",
    TRUE ~ "Not Significant"
  ))

# Create the volcano plot with facets for each method
volcano_plot_combined <- ggplot(combined_df, aes(x = avg_log2FC, y = -log10(p_val_adj), color = color)) +
  geom_point(alpha = 0.6, size = 2) +
  scale_color_manual(values = c("up" = "red", 
                                "down" = "blue", 
                                "Not Significant" = "gray")) +
  facet_wrap(~ method, scales = "free") +
  labs(title = "Volcano Plots Across Methods",
       x = "Log2 Fold Change",
       y = "-Log10 Adjusted P-value") +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 16),
    legend.position = "right"
  )

# Display the plot
print(volcano_plot_combined)


```

```{r}
sig_list_DESeq2 <- rownames(sig_DESeq2)
sig_list_bimod  <- rownames(sig_bimod)
sig_list_LR     <- rownames(sig_LR)
sig_list_MAST   <- rownames(sig_MAST)
sig_list_t      <- rownames(sig_t)
sig_list_wilcox <- rownames(sig_wilcox)

print(paste0("DESeq2 significant biomarkers: ", length(sig_list_DESeq2)))
print(paste0("bimod significant biomarkers: ", length(sig_list_bimod)))
print(paste0("LR significant biomarkers: ", length(sig_list_LR)))
print(paste0("MAST significant biomarkers: ", length(sig_list_MAST)))
print(paste0("t significant biomarkers: ", length(sig_list_t)))
print(paste0("wilcox significant biomarkers: ", length(sig_list_wilcox)))
print(paste0("common significant biomarkers: ", length(common_genes)))

```










-----------------------------------------------------------------------------------------------





 pseudo-bulk workflow


In Seurat, you can analyze pseudo bulk differential expression (DE) using DESeq2, which is commonly used for differential expression analysis in bulk RNA-seq data but can be adapted for single-cell data. This involves aggregating single-cell RNA-seq data into pseudo-bulk samples (by summing or averaging gene expression across cells in each condition) and then applying DESeq2 for differential expression analysis.

Benefits:
This approach leverages DESeq2's well-established statistical framework for differential expression, applying it to aggregated single-cell data.


Considerations:
Pseudo bulk analysis involves aggregating data, which may smooth out some single-cell-specific features. It’s essential to consider whether this approach aligns with your experimental design and research questions.


```{r}
library(DESeq2)

```



```{r}
seu.filtered$samples <- paste0(seu.filtered$stim, seu.filtered$ind)

head(seu.filtered@meta.data)

```



```{r}
cts <- AggregateExpression(seu.filtered, 
                    group.by = c("cell", "samples"),
                    assays = 'originalexp',
                    slot = "counts",
                    return.seurat = FALSE)


```


```{r}
cts <- cts$originalexp

```


```{r}
# transpose
cts.t <- t(cts)

```


```{r}
# convert to data.frame
cts.t <- as.data.frame(cts.t)

```

```{r}

head(seu.filtered@meta.data)

```


```{r}
# get values where to split
splitRows <- gsub('_.*', '', rownames(cts.t))


```



```{r}
# split data.frame
cts.split <- split.data.frame(cts.t,
                 f = factor(splitRows))


```


```{r}
cts.split.modified <- lapply(cts.split, function(x){
  rownames(x) <- gsub('.*_(.*)', '\\1', rownames(x))
  t(x)
  
})
dim(cts.split.modified$`CD14+ Monocytes`)

```
 
 DE analysis with CD14 Monocytes
 
```{r}

#  Get counts matrix

counts_CD14_Monocytes <- cts.split.modified$`CD14+ Monocytes`
```

```{r}
dim(counts_CD14_Monocytes)
```


```{r}
#  generate sample level metadata

colData <- data.frame(samples = colnames(counts_CD14_Monocytes))
```


```{r}
#create the metadata

colData <- colData %>%
  mutate(condition = ifelse(grepl('stim', samples), 'Stimulated', 'Control')) %>%
  column_to_rownames(var = 'samples')

```


```{r}
# Create DESeq2 object  

dds <- DESeqDataSetFromMatrix(countData = counts_CD14_Monocytes,
                       colData = colData,
                       design = ~ condition)

```

```{r}
# filter
keep <- rowSums(counts(dds)) >=10
dds <- dds[keep,]

```


```{r}
# run DESeq2
dds <- DESeq(dds)
```


```{r}
resultsNames(dds)


```


```{r}
res <- results(dds, name = "condition_Stimulated_vs_Control")
head(res , n= 20)
```



```{r}

# Create a new column to indicate the direction of exp
res$direction = ifelse(res$log2FoldChange > 0, "Up", "Down")
```


```{r}
library(ggplot2)
library(ggrepel)  

# VOLCANO Plot
ggplot(res, aes(x = log2FoldChange, y = -log10(padj), color = direction)) + 
  geom_point(alpha = 0.7, size = 2, shape = 21, stroke = 0.5) +  # Adjust appearance of points
    geom_text(aes(label = rownames(res)), check_overlap = TRUE, vjust = 3 , size =1.5)+
   
  geom_vline(xintercept = c(2, -2), linetype = "dashed", color = "grey50") +  # Vertical lines
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "grey50") +  # Horizontal line
  scale_color_manual(values = c("Up" = "blue", "Down" = "red")) + 
  scale_size_continuous(range = c(2, 6)) +  # Adjust size range for points
  theme_classic(base_size = 15) + 
  theme(
    legend.position = "bottom", 
    legend.title = element_text(face = "bold"),
    legend.text = element_text(size = 12),
    axis.title = element_text(face = "bold", size = 14),
    axis.text = element_text(size = 12),
   
  ) +
  labs(
    x = "Log2 Fold Change", 
    y = "-log10(Adjusted p-value)", 
    color = "Direction",
    size = "Significance",
    title = "Volcano Plot"
  )

```




```{r}
cell_plot <- DimPlot(seu.filtered, reduction = 'umap', group.by = 'cell', label = TRUE)
cond_plot <- DimPlot(seu.filtered, reduction = 'umap', group.by = 'stim')

cell_plot|cond_plot
```


```{r}

FeaturePlot(seu.filtered, features = 'IL1RN', min.cutoff = 'q10' , label = T)

```


```{r}

# Convert DESeqResults object to a data.frame
res_df <- as.data.frame(res)

# Filter, sort, and select top 10 up features
filtered_res <- res_df %>%
  filter(direction =="Up" & log2FoldChange > 2 ) %>%
  arrange(padj) %>%
  slice_head(n = 10)

# Extract row names from the filtered data frame
top_features <- rownames(res)[rownames(res) %in% rownames(filtered_res)]

# Print the top features
print(top_features)

```


```{r}
DotPlot(
  seu.filtered,
  top_features,
  assay = NULL,
  cols = c("lightgrey", "blue"),
  col.min = -2.5,
  col.max = 2.5,
  dot.min = 0,
  dot.scale = 6,
  idents = NULL,
  group.by = NULL,
  split.by = NULL,
  cluster.idents = FALSE,
  scale = TRUE,
  scale.by = "radius",
  scale.min = NA,
  scale.max = NA
) + theme(axis.text.x = element_text(angle = 90, hjust = 1 , size = 5))


```

```{r}
sig_list_bulk <- res_df %>%
  filter(padj < 0.05 & log2FoldChange > 2) %>%
  row.names

print(paste0('significant biomarkers in bulk: ',length(sig_list_bulk)))


```

```{r}
library(ggVennDiagram)

# Create a list of sets
sets <- list(
    wilcox = sig_list_wilcox,
    bimod  = sig_list_bimod,
    t      = sig_list_t,
    LR     = sig_list_LR,
    MAST   = sig_list_MAST,
    DESeq2 = sig_list_DESeq2,
    bulk   = sig_list_bulk
)

# Draw the Venn diagram with enhanced aesthetics
venn_plot <- ggVennDiagram(sets) + 
  scale_fill_gradient(low = "white", high = "#0073C2FF") +  # Adjust fill gradient
  labs(title = "Gene Overlap Across Different methods with bulk DEGs") +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),  # Center title
    legend.position = "right",  # Move legend to the right
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14),
    text = element_text(size = 14)
  )

venn_plot

```



```{r}
common_genes_with_bulk <- Reduce(intersect, rownames(sig_genes_list) , sig_list_bulk)
```


```{r}
names(res_df) <- paste0(names(res_df), ".bulk")
res_df$gene <- rownames(res_df)

names(DEGs_bimod) <- paste0(names(DEGs_bimod), ".bimod")
DEGs_bimod$gene <- rownames(DEGs_bimod)

names(DEGs_DESeq2) <- paste0(names(DEGs_DESeq2), ".DESeq2")
DEGs_DESeq2$gene <- rownames(DEGs_DESeq2)

names(DEGs_LR) <- paste0(names(DEGs_LR), ".LR")
DEGs_LR$gene <- rownames(DEGs_LR)

names(DEGs_MAST) <- paste0(names(DEGs_MAST), ".MAST")
DEGs_MAST$gene <- rownames(DEGs_MAST)

names(DEGs_t) <- paste0(names(DEGs_t), ".t")
DEGs_t$gene <- rownames(DEGs_t)

names(DEGs_wilcox) <- paste0(names(DEGs_wilcox), ".wilcox")
DEGs_wilcox$gene <- rownames(DEGs_wilcox)

merge_dat <- Reduce(function(x, y) merge(x, y, by = "gene"), 
                    list(DEGs_DESeq2, DEGs_bimod, DEGs_LR, DEGs_MAST, DEGs_t, DEGs_wilcox, res_df))
merge_dat <- merge_dat[order(merge_dat$pvalue.bulk), ]
```


```{r}
Idents(seu.filtered) <- "celltype.cnd"

```

```{r}
print(merge_dat[merge_dat$gene%in%common_genes[1:2],c('gene','p_val_adj.DESeq2','p_val_adj.MAST')])

```


```{r}
VlnPlot(seu.filtered, features = common_genes[1:2], idents = c("CD14+ Monocytes_stim", "CD14+ Monocytes_ctrl"), group.by = "stim") 

```


```{r}
VlnPlot(seu.filtered, features = common_genes_with_bulk[1:2], idents = c("CD14+ Monocytes_stim", "CD14+ Monocytes_ctrl"), group.by = "samples", ncol = 1) 
```


```{r}
# Filter for genes with padj < 0.05 in the specific method and padj > 0.05 in all other methods
only_wilcox <- merge_dat$gene[which(
  merge_dat$padj.bulk > 0.05 &
  merge_dat$p_val_adj.bimod > 0.05 &
  merge_dat$p_val_adj.DESeq2 > 0.05 &
  merge_dat$p_val_adj.t > 0.05 &
  merge_dat$p_val_adj.LR > 0.05 &
  merge_dat$p_val_adj.MAST > 0.05 &
  merge_dat$p_val_adj.wilcox < 0.05
)]

only_bimod <- merge_dat$gene[which(
  merge_dat$padj.bulk > 0.05 &
  merge_dat$p_val_adj.bimod < 0.05 &
  merge_dat$p_val_adj.DESeq2 > 0.05 &
  merge_dat$p_val_adj.t > 0.05 &
  merge_dat$p_val_adj.LR > 0.05 &
  merge_dat$p_val_adj.MAST > 0.05 &
  merge_dat$p_val_adj.wilcox > 0.05
)]

only_DESeq2 <- merge_dat$gene[which(
  merge_dat$padj.bulk > 0.05 &
  merge_dat$p_val_adj.bimod > 0.05 &
  merge_dat$p_val_adj.DESeq2 < 0.05 &
  merge_dat$p_val_adj.t > 0.05 &
  merge_dat$p_val_adj.LR > 0.05 &
  merge_dat$p_val_adj.MAST > 0.05 &
  merge_dat$p_val_adj.wilcox > 0.05
)]

only_LR <- merge_dat$gene[which(
  merge_dat$padj.bulk > 0.05 &
  merge_dat$p_val_adj.bimod > 0.05 &
  merge_dat$p_val_adj.DESeq2 > 0.05 &
  merge_dat$p_val_adj.t > 0.05 &
  merge_dat$p_val_adj.LR < 0.05 &
  merge_dat$p_val_adj.MAST > 0.05 &
  merge_dat$p_val_adj.wilcox > 0.05
)]

only_MAST <- merge_dat$gene[which(
  merge_dat$padj.bulk > 0.05 &
  merge_dat$p_val_adj.bimod > 0.05 &
  merge_dat$p_val_adj.DESeq2 > 0.05 &
  merge_dat$p_val_adj.t > 0.05 &
  merge_dat$p_val_adj.LR > 0.05 &
  merge_dat$p_val_adj.MAST < 0.05 &
  merge_dat$p_val_adj.wilcox > 0.05
)]

only_t <- merge_dat$gene[which(
  merge_dat$padj.bulk > 0.05 &
  merge_dat$p_val_adj.bimod > 0.05 &
  merge_dat$p_val_adj.DESeq2 > 0.05 &
  merge_dat$p_val_adj.t < 0.05 &
  merge_dat$p_val_adj.LR > 0.05 &
  merge_dat$p_val_adj.MAST > 0.05 &
  merge_dat$p_val_adj.wilcox > 0.05
)]

only_bulk <- merge_dat$gene[which(
  merge_dat$padj.bulk < 0.05 &
  merge_dat$p_val_adj.bimod > 0.05 &
  merge_dat$p_val_adj.DESeq2 > 0.05 &
  merge_dat$p_val_adj.t > 0.05 &
  merge_dat$p_val_adj.LR > 0.05 &
  merge_dat$p_val_adj.MAST > 0.05 &
  merge_dat$p_val_adj.wilcox > 0.05
)]

# Create a list of filtered genes
exclusive_genes_list <- list(
  "Wilcox" = only_wilcox,
  "Bimod" = only_bimod,
  "DESeq2" = only_DESeq2,
  "LR" = only_LR,
  "MAST" = only_MAST,
  "T" = only_t,
  "bulk" = only_bulk
)

# Print the number of filtered genes for each method
sapply(exclusive_genes_list, length)

```
```{r}
head(exclusive_genes_list)
```




```{r}
print(merge_dat[merge_dat$gene%in%c('CLTA','CAPG'),c('gene','p_val_adj.MAST',"p_val_adj.DESeq2", "p_val_adj.bimod", "p_val_adj.LR","p_val_adj.t", "p_val_adj.wilcox" ,'padj.bulk')])
```



```{r}
VlnPlot(seu.filtered, features <- c('CLTA','CAPG'), idents = c("CD14+ Monocytes_stim", "CD14+ Monocytes_ctrl"), group.by = "stim") 
```


```{r}
VlnPlot(seu.filtered, features <- c('CLTA','CAPG'), idents = c("CD14+ Monocytes_stim", "CD14+ Monocytes_ctrl"), group.by = "samples", ncol = 1) 
```



GRHPR is only expressed in LR method

```{r}
VlnPlot(seu.filtered, features <- c('GRHPR'), idents = c("CD14+ Monocytes_stim", "CD14+ Monocytes_ctrl"), group.by = "samples", ncol = 1) 
```

COPS7A is Only expressed in bimod method

```{r}
VlnPlot(seu.filtered, features <- c('COPS7A'), idents = c("CD14+ Monocytes_stim", "CD14+ Monocytes_ctrl"), group.by = "samples", ncol = 1) 
```


```{r}
    top_50_common    <- merge_dat[merge_dat[ , 1] %in% common_genes_with_bulk , ]
    top_50_common    <- top_50_common[1:50, ]

top_genes_common_mat <- counts_CD14_Monocytes[ rownames(counts_CD14_Monocytes) %in% top_50_common$gene , ]

```


```{r}
library(ComplexHeatmap)
set.seed(16863)
mycolors= rep("navy",dim(colData)[1])
mycolors[which(colData=="stimulated")]="green"

column_ha = HeatmapAnnotation(Sample_type= colData$condition)
Heatmap(top_genes_common_mat, name = "top", top_annotation = column_ha)
Heatmap(top_genes_common_mat, name = "top", top_annotation = column_ha, row_names_gp = gpar(fontsize = 5), column_names_gp = gpar(fontsize = 5, mycolors))

```

```{r}
DotPlot(
  seu.filtered,
  common_genes_with_bulk,
  assay = NULL,
  cols = c("lightgrey", "blue"),
  col.min = -2.5,
  col.max = 2.5,
  dot.min = 0,
  dot.scale = 6,
  idents = NULL,
  group.by = NULL,
  split.by = NULL,
  cluster.idents = FALSE,
  scale = TRUE,
  scale.by = "radius",
  scale.min = NA,
  scale.max = NA
) + theme(axis.text.x = element_text(angle = 90, hjust = 1 , size = 5))


```



```{r}
print(merge_dat[merge_dat$gene%in%c('CCL8' ,'ISG15' , 'IFI6' ),c('gene','p_val_adj.MAST',"p_val_adj.DESeq2", "p_val_adj.bimod", "p_val_adj.LR","p_val_adj.t", "p_val_adj.wilcox" ,'padj.bulk')])
```



```{r}

FeaturePlot(seu.filtered, features = c("CCL8" ,'ISG15' , 'IFI6') , label = T )

```




```{r}
VlnPlot(seu.filtered, features <- c('CCL8' ,'ISG15' ,'IFI6'), idents = c("CD14+ Monocytes_stim", "CD14+ Monocytes_ctrl"), group.by = "samples", ncol = 1) 
```


```{r}
```




